Gravitational wave sources in our Galactic backyard: Predictions for BH and NS binaries in LISA

### 2021-07-10 paper/tex/sections/results.tex
update text to match new properties grid

The reasoning for these unequal mass ratio systems is as follows: in order to produce a BHBH, most formation channels require at least the first mass transfer to be stable. This stability is strongly dependent on the mass ratio such that equal mass ratios (at the moment of mass transfer) are preferred for creating BHBHs. Yet, since stellar winds are so strong at high metallicity, and even stronger for more massive stars, the primary star will experience significant mass loss and so an initially \textit{unequal} mass ratio is preferred so that the masses are more balanced at the first instance of mass transfer. Since mass transfer occurs after the end of the main sequence for most of our BHBHs, the star will have a well defined core and these core masses, which go on to form BHs, will reflect the initially unequal mass ratios.

\subsubsection{Neutron Star Mass}
The neutron star mass distribution shows that most neutron stars have low masses, with $77\%$ and $91\%$ having masses below $1.7 \unit{M_{\odot}}$ for BHNSs and NSNSs respectively. The lack of neutron stars around $1.7 \unit{M_{\odot}}$ and the subsequent small peaks are a result of the discontinuous nature of the \citet{Fryer+2012} remnant mass prescription but do not have strong effects on our results since the vast majority of NSs are formed with lower masses.

Both distributions have a notable peak around $1.26 \unit{M_{\odot}}$, which is a result of the combination of two effects. Firstly, we set the remnant mass for all electron-capture supernovae to $1.26 \unit{M_\odot}$ (see Sec.~\ref{sec:fiducial_physics}) and thus this leads to a pileup when many systems are formed through ECSN. In addition, the Fryer remnant mass prescription gives a fixed fallback mass for any star with a CO core mass less than $2.5 \unit{M_\odot}$, such that many NSs are given the identical mass of $1.278 \unit{M_\odot}$ in the \textit{delayed} prescription \citep[see][Eq.~19]{Fryer+2012}.

@@

\subsubsection{Luminosity Distance}
Each DCO's luminosity distance distribution peaks around $8 \unit{kpc}$ since this is the distance to the centre of the Milky Way and thus the most dense location of DCOs. There is a bias in each distribution for systems at lower distances since closer binaries are easier to detect. This bias is most prominent for the NSNS distribution since, on average, their lower relative masses require a smaller distance in order to be detected.

However, more surprisingly, we also see that the NSNS distribution extends to higher distances than the other DCOs. The reason for this is that the NSNS population has the highest fraction of ``mildly'' eccentric systems ($0.01 < e < 0.3$). The BHNS population has a much higher fraction of effectively circular systems ($e < 0.01$), which emit weaker gravitational wave compared to equivalent eccentric systems. Therefore, despite their relatively higher masses, the maximum distance at which a source is detectable is generally lower than the eccentric NSNSs and hence the distribution tends to zero at smaller distances. Conversely, the BHBH population has a higher fraction of \textit{high} eccentricity systems $e > 0.3$. Although one may naively expect that this would result in stronger signals (and so further distances), for a system to have these high eccentricities in LISA, it must still be early in its evolution and thus have a low orbital frequencies. The result of this is that high eccentricity systems tend to have lower SNRs and so cannot be detected at large distances. Overall we see that the eccentricity distribution of NSNSs occupies a ``sweet spot'' where the gravitational wave power is increased compared to circular systems, but it isn't too high that the frequency is significantly impacted. This means that NSNSs can be seen out to the largest distances of the three DCO types.

\subsubsection{Lookback Time}
Although when creating each mock Milky Way galaxy the lookback time is drawn independently from other parameters, and in the same way for every DCO type, the distributions are clearly different for each type of DCO. This is because for a binary to be detectable, it must have an orbital frequency within the LISA band. Therefore a binary needs to have completed most of its inspiral to be visible in the LISA band and thus detectable systems will tend to have lookback time that are close to their merger times, which \textit{are} a function of other binary parameters.

Since most of the DCOs are formed through common envelope events, their initial separations are relatively tight and so if a binary it given a long lookback time, it will have merged before the LISA mission. This explains the peak in each distribution at short lookback times. The later lookback times correspond to systems that are formed through channels in which mass transfer is only stable and no common envelope event occurs.

The trends across the different DCO types can be explained by considering that the merger time is a function of the masses, frequency and eccentricity of th source. This dependence can be approximately written as $t_{\rm merge} \propto f_{\rm orb}^{-4} m_1^{-3} (1 - e^2)^{-7/2}$,  \citep[][Eq.~5.14]{Peters+1964}. Therefore, although NSNSs are the lowest mass systems, their relatively higher orbital frequencies means that they have the shortest merger times. This same logic implies that BHNSs should have the next shortest merger times and then BHBHs. This order is flipped due to the fact that high eccentricity results in a shorter merger time and we find that LISA detectable BHNSs are mostly circular, whilst BHBHs have a significant highly eccentric subpopulation.

>>>>

\subsubsection{Mass Ratio}
The mass ratio distribution for each DCO type is relatively distinct from the others. The majority of NSNSs have a mass ratio close to unity, with 90\% of systems having $q > 0.8$. The reason for the concentration around equal masses is that most NSs are formed either through electron-capture supernovae or from low mass stars. We set the remnant mass for any system formed through ECSN to $1.26 \unit{M_{\odot}}$ (see Sec.~\ref{sec:fiducial_physics}). The remnant mass prescription that we use gives a fixed fallback mass for any star with a CO core mass less than $2.5 \unit{M_\odot}$, such that many NSs are given the identical mass of $1.278 \unit{M_\odot}$ in the \textit{delayed} prescription \citep[see][Eq.~19]{Fryer+2012}. This means that many NSs are formed with equal masses and hence we see a mass ratio distribution peaked around unity.

In contrast, only 9\% of detectable BHBHs are formed with $q > 0.8$ and distribution peaks around $q = 0.4$. The reasoning for these unequal mass ratio systems is as follows: in order to produce a BHBH, most formation channels require at least the first mass transfer to be stable. This stability is strongly dependent on the mass ratio such that equal mass ratios (at the moment of mass transfer) are preferred for creating BHBHs. Yet, since stellar winds are so strong at high metallicity, and even stronger for more massive stars, the primary star will experience significant mass loss and so an initially \textit{unequal} mass ratio is preferred so that the masses are more balanced at the first instance of mass transfer. Since mass transfer occurs after the end of the main sequence for most of our BHBHs, the star will have a well defined core and these core masses, which go on to form BHs, will reflect the initially unequal mass ratios.

We find that detectable BHNSs have even more unequal mass ratios, such that the distribution is almost entire disjoint from that of NSNSs. Moreover, the mass ratio distribution is bimodal, where the two peaks arise from two distinct formation scenarios. Around two thirds of detectable BHNSs experience at least one common envelope event, whilst the last third are formed through only stable mass transfer. The first peak at $q = 0.18$ is from systems that experience at least one CE and occurs at the expected mass ratio, which approximately follows the mean BH mass ($\sim 6.5 \unit{M_\odot}$) and NS mass ($\sim 1.2 \unit{M_\odot}$). Yet we also see a second peak at higher mass ratios around $q = 0.34$, which arise from the fraction of the population that underwent only stable mass transfer. The stability of mass transfer is strongly dependent on the mass ratio and thus these systems have a bias for more equal mass ratio systems, leading to a peak at higher $q$.

@@

\subsubsection{Time since formation}
The progenitors of Galactic DCOs in our sample are formed throughout the history of the Milky Way, with more formed at earlier times (see Fig.~\ref{fig:galaxy_schematic}). In contrast, we find that the \textit{detectable} systems are primarily formed recently, with the majority formed in the past $2 \unit{Gyr}$, in addition to a tail of systems out to earlier times.

The peak at recent times is due to the fact that most binaries in our sample are formed with merger times on the order of a couple of Gyr (since most form through a common envelope phase that significantly tightens the binary). For a binary to be in the LISA band and detectable it must be near the end of its inspiral and so LISA will be most sensitive to sources that were formed a couple of Gyr ago.

The peak is sharpest for NSNSs as they have the highest orbital frequencies and largest fraction of eccentric system, which result in shorter inspiral times and therefore later formation times. Conversely, the peak is less prominent for BHNSs as they are the most circular, have moderate orbital frequencies and masses.

\subsubsection{Time until merger}
The final panel of Fig.~\ref{fig:fiducial_pdf_distributions} shows the remaining time until merger for each of the DCO types at the start of LISA mission. The distributions are strikingly similar and peak with merger times of around a Myr.

The merger time is a function of the mass, frequency and eccentricity of the sources, such that more massive, higher frequency and more eccentric sources merge faster \citep[][Eq.~5.14]{Peters+1964}. So, despite the fact that each DCO type often has higher values in any one of these properties, the convolution of all three tends to negate the differences. For example, NSNSs have the highest orbital frequencies and are mildly eccentric whilst BHNSs have moderate orbital frequencies and are more circular. However, BHNSs are more massive in general and so the overall merger times are distributed very similarly for both DCO types.
